---
title: "05_visualization"
author: "Evan Muise"
date: "20/10/2021"
output: html_document
---


```{r}
library(tidyverse) # in use
library(tidymodels) # t tests
library(cowplot) # plot additions
library(patchwork) # plot additions
library(ggrepel) # for t-test plot
library(ggridges) # for ridge plot
library(magick) # for image cropping of 3d vectors
#library(gtsummary)
#library(ggpubr)
library(flextable)
library(patchwork)

library(sf) # in use
 # in use for data

library(ggsci) # in use for aaas
library(scico) # in use for ridge plot
#library(ggthemes) 

library(plot3D) # in use for 3d plot
library(plot3Drgl) # not in use
library(gridGraphics) # probably in use

library(raster)
library(officer)
library(RColorBrewer)
library(terra)
library(ggnewscale)

my_theme <- theme_bw() +
  theme(panel.grid = element_blank())
theme_set(my_theme)
set.seed(69420)

elev_agg_val <- 50

```

```{r}
load(here::here("data", "backup.RData"))
source(here::here("scripts", "get_keys.R"))
vlce_colours <- keys$vlce %>% pull(ntems_colour)
```

Elevation & Protected/Unprotected (BEC) Figures

```{r elev-bec}
elev_zone_data <- elev_nw %>% 
  ungroup() %>%
  filter(elev < 32000) %>%
  separate(subzone, into = c("zone", "subzone")) %>% 
  group_by(protected, elev, zone) %>% 
  summarize(n = sum(count)) %>% 
  mutate(perc = n / sum(n)) 

elev_agg <- elev_zone_data %>%
  mutate(elev = elev %/% elev_agg_val * elev_agg_val) %>%
  group_by(protected, elev, zone) %>%
  summarize(n = sum(n)) %>%
  mutate(perc = n / sum(n))

elev_zone_plot_agg <- elev_agg %>%
  ggplot(aes(x = elev, y = perc, fill = zone)) +
  geom_area(stat = "identity") +
  facet_wrap(~str_to_title(protected)) +
  coord_flip() +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Elevation (m)",
       y = "% Cover",
       fill = "BEC Zone") +
  theme(legend.position = "left") +
  scale_fill_igv()

protected_elev <- elev_zone_data %>%
  ungroup() %>%
  pivot_wider(values_from = n, names_from = protected) %>% 
  group_by(elev) %>%
  summarize(protected = sum(protected, na.rm = T),
            unprotected = sum(unprotected, na.rm = T)) %>%
  mutate(total = protected + unprotected,
         per_prot = protected / total,
         per_unprot = 1 - per_prot) %>%
  pivot_longer(per_prot:per_unprot, names_to = "value", values_to = "percent")

prot_elev_plot_agg <- protected_elev %>% 
  mutate(elev = elev %/% elev_agg_val * elev_agg_val) %>% 
  group_by(elev) %>%
  summarize(protected = sum(protected),
            unprotected = sum(unprotected)) %>%
  mutate(total = protected + unprotected,
         per_prot = protected / total,
         per_unprot = 1 - per_prot) %>%
  ggplot(aes(x = elev, y = per_prot)) +
  #geom_smooth() +
  #geom_point() +
  geom_area() +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
  labs(x = NULL,
       y = "% Protected",
       fill = NULL) +
  theme(axis.text.y = element_blank(),
        legend.position = c(0.7, 0.85)) +
  coord_flip()

mosaic_elev_plot <- plot_grid(elev_zone_plot_agg, prot_elev_plot_agg,
          rel_widths = c(3, 1),
          align = 'h', axis = 'tb')

ggsave(here::here("outputs", "bec_elev.png"), plot = mosaic_elev_plot, 
       device = "png",
       height = 5.5,
       width = 8.5)
```

Updated elev bec w/ hist

```{r elev-bec-hist}
for_pyramid <- elev_agg %>%
  filter(elev < 32000) %>%
  dplyr::select(!perc) %>%
  group_by(protected, elev) %>%
  mutate(perc = n / sum(n),
         ha = 30 * 30 * n / 10000 / 1e6) %>%
  left_join(keys$bec)

prot_line <- for_pyramid %>%
  summarize(n = sum(n)) %>%
  group_by(elev) %>%
  mutate(perc = n / sum(n)) %>%
  filter(protected == "protected") 

hist_prot <- for_pyramid %>%
  filter(protected == "protected") %>%
  ggplot(aes(x = elev, weight = ha)) +
  #geom_histogram(data = subset(for_pyramid, protected == "unprotected")) +
  geom_histogram(aes(weight = ha * (-1)), fill = "black", bins = 20) +
  #facet_wrap(~protected) +
  coord_flip() +
  labs(x = "Elevation (m)",
       y = "Area (million ha)") +
  scale_y_continuous(limits = c(-15, 0),
                     breaks = c(-15, -10, -5, 0),
                     labels = c(15, 10, 5, 0)) +
  facet_wrap(~str_to_title(protected)) +
  scale_x_continuous(limits = c(0, 4600))

hist_unprot <- for_pyramid %>%
  filter(protected == "unprotected") %>%
  ggplot(aes(x = elev, weight = ha), fill = "black") +
  geom_histogram(data = subset(for_pyramid, protected == "unprotected"), fill = "black", bins = 20) +
  #geom_histogram(data = subset(for_pyramid, protected == "protected"), aes(weight = n * (-1))) +
  #facet_wrap(~protected) +
  coord_flip() +
  labs(x = NULL,
       y = "Area (million ha)") +
  theme(axis.text.y = element_blank()) +
  scale_y_continuous(limits = c(0, 15)) +
  facet_wrap(~str_to_title(protected)) +
  scale_x_continuous(limits = c(0, 4600))


area_prot <- for_pyramid %>%
  filter(protected == "protected") %>%
  ggplot(aes(x = elev, y = perc)) +
  geom_area(aes(fill = fct_reorder(zone_nm, order))) +
  #geom_line(data = prot_line, lty = "dashed") +
  coord_flip() +
  scale_fill_igv() +
  #geom_hline(yintercept = 0, lty = "dashed", colour = "red") +
  scale_y_continuous(labels = scales::percent) +
  labs(x = NULL,
       y = "Proportion of Landscape",
       fill = NULL) +
  theme(legend.position = "bottom",
        axis.text.y = element_blank()) +
  facet_wrap(~str_to_title(protected)) +
  scale_x_continuous(limits = c(0, 4600))

area_unprot <- for_pyramid %>%
  filter(protected == "unprotected") %>%
  ggplot(aes(x = elev, y = perc, fill = fct_reorder(zone_nm, order))) +
  geom_area() +
  coord_flip() +
  scale_fill_igv() +
  #geom_hline(yintercept = 0, lty = "dashed", colour = "red") +
  scale_y_continuous(labels = scales::percent) +
  labs(x = NULL,
       y = "Proportion of Landscape",
       fill = NULL) +
  theme(legend.position = "bottom",
        axis.text.y = element_blank()) +
  facet_wrap(~str_to_title(protected)) +
  scale_x_continuous(limits = c(0, 4600))

bec_elev_hist <- (hist_prot + area_prot + area_unprot + hist_unprot) +
  plot_annotation(tag_levels = "a") +
  plot_layout(nrow = 1, 
              guides = "collect",
              widths = c(1, 3, 3, 1)) &
  theme(legend.position = "bottom")

ggsave(here::here("outputs", "bec_elev_hist.png"), bec_elev_hist, device = "png",
       height = 6.5, width = 10)
```

```{r}
for_pyramid %>%
  dplyr::select(protected:n) %>%
  pivot_wider(names_from = protected,
              values_from = n) %>%
  mutate(protected = ifelse(is.na(protected), 0, protected),
         unprotected = ifelse(is.na(unprotected), 0, unprotected),
         ratio = (protected - unprotected) / (protected + unprotected)) %>%
  filter(!is.infinite(ratio) |
           ratio > 5000) %>%
  ggplot(aes(x = elev, y = ratio, colour = zone)) +
  geom_line() +
  scale_colour_igv() +
  coord_flip()


```

Elevation & Protected/Unprotected (VLCE) Figures

```{r elev-vlce}
elev_lcc_nw <- elev_w %>%
  filter(elev < 32000) %>%
  mutate(elev = elev %/% elev_agg_val * elev_agg_val) %>%
  group_by(protected, elev, class_name) %>%
  summarize(n = sum(count)) %>%
  mutate(perc = n / sum(n)) %>%
  left_join(keys$vlce) %>%
  mutate(name_clean = fct_rev(fct_reorder(name_clean, class_val)))

elev_lcc_agg <- elev_lcc_nw %>%
  ggplot(aes(x = elev, y = perc, fill = name_clean)) +
  geom_area(stat = "identity") +
  coord_flip() +
  facet_wrap(~str_to_title(protected)) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = vlce_colours) +
  labs(x = "Elevation (m)",
       y = "% Cover",
       fill = "Land Cover") +
  theme(legend.position = "left")

mosaic_lcc_plot <- plot_grid(elev_lcc_agg, prot_elev_plot_agg,
          rel_widths = c(3, 1),
          align = 'h', axis = 'tb')

ggsave(here::here("outputs", "lcc_elev.png"), plot = mosaic_lcc_plot, device = "png")
```

Updated VLCE Elev w/ Histogram

```{r elev_vlce_hist}
for_pyramid_vlce <- elev_lcc_nw %>%
  dplyr::select(protected, elev, name_clean, n) %>%
  group_by(protected, elev) %>%
  mutate(perc = n / sum(n),
         ha = 30 * 30 * n / 10000 / 1e6)

hist_prot <- for_pyramid_vlce %>%
  filter(protected == "protected") %>%
  ggplot(aes(x = elev, weight = ha)) +
  #geom_histogram(data = subset(for_pyramid, protected == "unprotected")) +
  geom_histogram(aes(weight = ha * (-1)), fill = "black", bins = 20) +
  #facet_wrap(~protected) +
  coord_flip() +
  labs(x = "Elevation (m)",
       y = "Area (million ha)") +
  scale_y_continuous(limits = c(-15, 0),
                     breaks = c(-15, -10, -5, 0),
                     labels = c(15, 10, 5, 0)) +
  facet_wrap(~str_to_title(protected)) +
  scale_x_continuous(limits = c(0, 4600))

hist_unprot <- for_pyramid_vlce %>%
  filter(protected == "unprotected") %>%
  ggplot(aes(x = elev, weight = ha), fill = "black") +
  geom_histogram(fill = "black", bins = 20) +
  #geom_histogram(data = subset(for_pyramid, protected == "protected"), aes(weight = n * (-1))) +
  #facet_wrap(~protected) +
  coord_flip() +
  labs(x = NULL,
       y = "Area (million ha)") +
  theme(axis.text.y = element_blank()) +
  scale_y_continuous(limits = c(0, 15)) +
  facet_wrap(~str_to_title(protected)) +
  scale_x_continuous(limits = c(0, 4600))


area_prot <- for_pyramid_vlce %>%
  filter(protected == "protected") %>%
  ggplot(aes(x = elev, y = perc, fill = fct_rev(name_clean))) +
  geom_area() +
  coord_flip() +
  scale_fill_manual(values = vlce_colours) +
  #geom_hline(yintercept = 0, lty = "dashed", colour = "red") +
  scale_y_continuous(labels = scales::percent) +
  labs(x = NULL,
       y = "Proportion of Landscape",
       fill = NULL) +
  theme(legend.position = "bottom",
        axis.text.y = element_blank()) +
  facet_wrap(~str_to_title(protected)) +
  scale_x_continuous(limits = c(0, 4600))

area_unprot <- for_pyramid_vlce %>%
  filter(protected == "unprotected") %>%
  ggplot(aes(x = elev, y = perc, fill = fct_rev(name_clean))) +
  geom_area() +
  coord_flip() +
  scale_fill_manual(values = vlce_colours) +
  #geom_hline(yintercept = 0, lty = "dashed", colour = "red") +
  scale_y_continuous(labels = scales::percent) +
  labs(x = NULL,
       y = "Proportion of Landscape",
       fill = NULL) +
  theme(legend.position = "bottom",
        axis.text.y = element_blank()) +
  facet_wrap(~str_to_title(protected)) +
  scale_x_continuous(limits = c(0, 4600))

vlce_elev_hist <- (hist_prot + area_prot + area_unprot + hist_unprot) +
  plot_annotation(tag_levels = "a") +
  plot_layout(nrow = 1, 
              guides = "collect",
              widths = c(1, 3, 3, 1)) &
  theme(legend.position = "bottom")

ggsave(here::here("outputs", "lcc_elev_hist.png"), vlce_elev_hist, device = "png",
       height = 6.5, width = 10)
```

```{r}
library(reldist)

splits <- elev_agg %>%
  group_by(protected, zone) %>%
  group_split()

my_quantile <- function(df) {
  zone = df %>% 
    pull(zone) %>%
    unique()
  
  protected = df %>% 
    pull(protected) %>%
    unique()
  
  outdf <- as.data.frame(wtd.quantile(df$elev, q = c(0.25, 0.5, 0.75), weight = df$n))
  outdf["q"] = rownames(outdf)
  colnames(outdf) <- c("elev", "q")
  
  tibble(outdf) %>%
    pivot_wider(names_from = q,
                values_from = elev) %>%
    janitor::clean_names() %>%
    mutate(iqr = x75_percent - x25_percent,
           min = ifelse(x25_percent - iqr > 0, x25_percent - iqr, 0),
           max = x75_percent + iqr) %>%
    dplyr::select(!iqr) %>%
    bind_cols(zone = zone, protected = protected)
}

boxplot_df <- map_dfr(splits, my_quantile)

lcc_splits <- elev_lcc_nw %>%
  dplyr::select(protected, elev, class_name, n) %>%
  group_by(protected, class_name) %>%
  group_split()

my_quantile <- function(df) {
  zone = df %>% 
    pull(class_name) %>%
    unique()
  
  protected = df %>% 
    pull(protected) %>%
    unique()
  
  outdf <- as.data.frame(wtd.quantile(df$elev, q = c(0.25, 0.5, 0.75), weight = df$n))
  outdf["q"] = rownames(outdf)
  colnames(outdf) <- c("elev", "q")
  
  tibble(outdf) %>%
    pivot_wider(names_from = q,
                values_from = elev) %>%
    janitor::clean_names() %>%
    mutate(iqr = x75_percent - x25_percent,
           min = ifelse(x25_percent - iqr > 0, x25_percent - iqr, 0),
           max = x75_percent + iqr) %>%
    dplyr::select(!iqr) %>%
    bind_cols(zone = zone, protected = protected)
}

lcc_boxplot <- map_dfr(lcc_splits, my_quantile)
```

```{r}
bec_elev_boxes <- boxplot_df %>%
  ggplot(aes(x = fct_reorder(zone, x50_percent))) +
  geom_boxplot(aes(
    lower = x25_percent,
    upper = x75_percent,
    middle = x50_percent,
    ymin = min,
    ymax = max,
    fill = str_to_title(protected)),
    stat = "identity") +
  labs(x = "BEC Zone",
       y = NULL,
       fill = NULL) +
  coord_flip() +
  theme(legend.position = "none",
        axis.text.x = element_blank())

lcc_elev_boxes <- lcc_boxplot %>%
  left_join(keys$vlce, by = c("zone" = "class_name")) %>%
  ggplot(aes(x = fct_reorder(name_clean, x50_percent))) +
  geom_boxplot(aes(
    lower = x25_percent,
    upper = x75_percent,
    middle = x50_percent,
    ymin = min,
    ymax = max,
    fill = str_to_title(protected)),
    stat = "identity") +
  labs(x = "Land Cover Class",
       y = "Elevation (m)",
       fill = NULL) +
  coord_flip() +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 15)) +
  theme(legend.position = "bottom",
        legend.justification = c(0, 1),
        legend.margin = margin(t = 0, r = 0, b = 0, l = -69, unit = "pt"))
  
```


```{r}
(elev_boxplots <- bec_elev_boxes / lcc_elev_boxes +
  plot_annotation(tag_levels = "a"))

ggsave(here::here("outputs", "elev_boxplots.png"), 
       plot = elev_boxplots, 
       device = "png",
       width = 3,
       height = 6.5,
       units = "in")
```

bec-lcc-intersection
```{r}
bec_lcc <- elev_w %>%
  filter(elev < 32000) %>%
  mutate(elev = elev %/% elev_agg_val * elev_agg_val) %>%
  separate(subzone, into = c("zone", "subzone")) %>%
  group_by(protected, zone, class_name, elev) %>%
  summarize(n = sum(count)) %>%
  summarize(mean_elev = weighted.mean(elev, n)) %>%
  pivot_wider(names_from = protected, values_from = mean_elev)

bec_lcc %>%
  ungroup() %>%
  complete(zone, class_name) %>%
  mutate(difference = protected - unprotected) %>%
  ggplot(aes(x = str_to_title(class_name), y = fct_rev(zone), fill = difference)) +
  geom_tile() +
  scale_fill_scico(palette = "vik", limits = c(-800, 800)) +
  labs(x = "Land Cover Class",
       y = "BEC Zone",
       fill = str_wrap("Elevation Difference", width = 10)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```




BEC proportions

```{r bec-prop}
bec <- vlce %>%
  separate(subzone, into = c("zone", "subzone")) %>%
  group_by(protected, zone) %>%
  summarize(count = sum(n, na.rm = T)) %>%
  mutate(perc = count / sum(count))

bec_prot_donut <- vlce %>%
  separate(subzone, into = c("zone", "subzone")) %>%
  group_by(zone, protected) %>%
  summarize(count = sum(n)) %>%
  mutate(perc = count / sum(count)) 

bec_conch_df <- bec_prot_donut %>%
  filter(protected == "protected") %>%
  left_join(keys$bec) %>%
  #left_join(bec_perc_terr, by = c("zone" = "zone")) %>%
  mutate(zone_nm = fct_reorder(zone_nm, perc),
         zone = fct_reorder(zone, perc, .fun = max)) 

write_csv(bec_conch_df, here::here("outputs", "bec_prop_protected.csv"))

bec_conch <- bec_conch_df %>%
  ggplot(aes(x = fct_reorder(zone, perc), y = perc, fill = zone_nm)) +
  geom_bar(stat = "identity") +
  geom_hline(aes(yintercept = 0.17, colour = "2020 Protection Target (Global)"), lty = "dotted", size = 1) +
  geom_hline(aes(yintercept = 0.25, colour = "2025 Protection Target (Canada)"), lty = "dotted", size = 1) +
  geom_hline(aes(yintercept = 0.154, colour = "2020 BC Proportion Protected"), lty = "dotted", size = 1) +
  geom_hline(aes(yintercept = 0.3), lty = "dotted", size = 1) +
  geom_text(aes(x = 1, y = 0.135), label = "15.4%") +
  geom_text(aes(x = 1, y = 0.19), label = "17%") +
  geom_text(aes(x = 1, y = 0.27), label = "25%", colour = "black") +
  geom_text(aes(x = 1, y = 0.32, label = "30%")) +
  scale_colour_aaas() +
  scale_fill_igv() +
  coord_polar() +
  theme_void() + 
  theme(axis.line = element_blank(),
        axis.title = element_blank(), 
        axis.ticks = element_blank(), 
        axis.text.y = element_blank(),
        #legend.position = "bottom",
        legend.box = "vertical") +
  labs(fill = NULL,
       colour = NULL)

ggsave(here::here("outputs", "bec_conch.png"), plot = bec_conch, device = "png",
       height = 5.5, width = 8.5)

bec_bar <- bec_conch_df %>%
  ggplot(aes(x = perc, y = fct_reorder(zone, perc), fill = zone)) +
  geom_bar(stat = "identity") +
  geom_vline(aes(xintercept = 0.17), colour = "#00BA38", lty = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0.25), colour = "#619CFF", lty = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0.154), colour = "#F8766D", lty = "dashed", size = 1) +
  geom_hline(aes(yintercept = 0.17, colour = "2020 Protection Target (Global)"), lty = "dashed", size = 1) +
  geom_hline(aes(yintercept = 0.25, colour = "2025 Protection Target (Canada)"), lty = "dashed", size = 1) +
  geom_hline(aes(yintercept = 0.154, colour = "2020 BC Proportion Protected"), lty = "dashed", size = 1) +
  guides(col = guide_legend(ncol = 1,
                            keyheight = 0.75,
                            default_unit = "inch",
                            vjust = 1)) +
  theme(legend.position = "bottom",
        legend.key.size = unit(2, "lines"),
        legend.justification = c(0, 1),
        legend.margin = margin(t = 0, r = 0, b = 0, l = -20, unit = "pt")) +
  scale_fill_igv(guide = "none") +
  #theme(legend.position = "none") +
  labs(x = "% Protected",
       y = NULL,
       colour = NULL) +
  scale_x_continuous(labels = scales::percent) +
  scale_y_discrete(labels = function(x) str_wrap(x, width = 10))


bec_bar

ggsave(here::here("outputs", "bec_bar.png"), plot = bec_bar, device = "png",
       height = 6, width = 3)
```

Cell to pull a legend for BEC zones for later on

```{r legend}
for_legend <- bec_conch_df %>% 
  ggplot(aes(x = zone, y = count, fill = zone_nm)) +
  geom_col() +
  scale_fill_igv() +
  labs(fill = NULL)


bec_legend <- cowplot::get_legend(for_legend + theme(legend.box.margin = margin(0, 0, 0, -200)))

ggsave(here::here("outputs", "bec_legend.png", bec_legend, device = "png"))
```


VLCE Conch

```{r vlce-conch}
vlce_conch_df <- vlce %>%
  group_by(vlce, protected) %>%
  summarize(count = sum(n, na.rm = T)) %>%
  mutate(perc = count / sum(count)) %>%
  filter(protected == "protected") %>%
  left_join(keys$vlce, by = c("vlce" = "class_val")) %>%
  dplyr::select(name_clean, protected, perc, count) %>%
  ungroup()

vlce_conch <- vlce_conch_df %>%
  ggplot(aes(x = fct_reorder(as.factor(vlce), perc), y = perc, fill = fct_reorder(name_clean, vlce, .fun = max))) +
  geom_bar(stat = "identity") +
  geom_hline(aes(yintercept = 0.17, colour = "2020 Protection Target (Global)"), lty = "dotted", size = 1) +
  geom_hline(aes(yintercept = 0.25, colour = "2025 Protection Target (Canada)"), lty = "dotted", size = 1) +
  geom_hline(aes(yintercept = 0.154, colour = "2020 BC Proportion Protected"), lty = "dotted", size = 1) +
  geom_hline(aes(yintercept = 0.33), lty = "dotted", size = 1) +
  geom_text(aes(x = 1, y = 0.135), label = "15.4%") +
  geom_text(aes(x = 1, y = 0.19), label = "17%") +
  geom_text(aes(x = 1, y = 0.27), label = "25%", colour = "black") +
  geom_text(aes(x = 1, y = 0.35, label = "33%")) +
  scale_colour_aaas() +
  scale_fill_manual(values = vlce_colours) +
  coord_polar() +
  theme_void() + 
  theme(axis.line = element_blank(),
        axis.title = element_blank(), 
        axis.ticks = element_blank(), 
        axis.text.y = element_blank(),
        #legend.position = "bottom",
        legend.box = "vertical") +
  labs(fill = NULL,
       colour = NULL) +
  guides(fill = guide_legend(order = 2),
         colour = guide_legend(order = 1))

ggsave(here::here("outputs", "vlce_conch.png"), plot = vlce_conch, device = "png",
       height = 5.5, width = 8.5)

vlce_bar <- vlce_conch_df %>%
  ggplot(aes(x = perc, y = fct_reorder(as.factor(name_clean), perc), fill = fct_reorder(name_clean, vlce, .fun = max))) +
  geom_bar(stat = "identity") +
  geom_vline(aes(xintercept = 0.17), colour = "#00BA38", lty = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0.25), colour = "#619CFF", lty = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0.154), colour = "#F8766D", lty = "dashed", size = 1) +
  geom_hline(aes(yintercept = 0.17, colour = "2020 Protection Target (Global)"), lty = "dashed", size = 1) +
  geom_hline(aes(yintercept = 0.25, colour = "2025 Protection Target (Canada)"), lty = "dashed", size = 1) +
  geom_hline(aes(yintercept = 0.154, colour = "2020 BC Proportion Protected"), lty = "dashed", size = 1) +
  guides(col = guide_legend(ncol = 1,
                            keyheight = 0.75,
                            default_unit = "inch",
                            vjust = 1)) +
  theme(legend.position = "bottom",
        legend.key.size = unit(2, "lines"),
        legend.justification = c(0, 1),
        legend.margin = margin(t = 0, r = 0, b = 0, l = -20, unit = "pt")) +
  scale_fill_manual(values = vlce_colours, guide = "none") +
  #theme(legend.position = "none") +
  labs(x = "% Protected",
       y = NULL,
       colour = NULL) +
  scale_x_continuous(labels = scales::percent) +
  scale_y_discrete(labels = function(x) str_wrap(x, width = 10))

vlce_bar

ggsave(here::here("outputs", "vlce_bar.png"), vlce_bar, device = "png", 
       height = 6, width = 3)
```

T-Test Plots

```{r t-tests}
t_tests_b <- t_tests %>%
  filter(!variable %in% c("elev", "slope")) %>%
  mutate(p.value = p.value * 496) ## bonferroni correction

t_tests_zonal <- t_tests_b %>%
  separate(subzone, into = c("zone", "subzone")) %>% 
  mutate(sig0.05 = ifelse(p.value < 0.05, T, F),
         sig0.01 = ifelse(p.value < 0.01, T, F)) %>%
  group_by(zone, variable) %>%
  summarize(sig0.05 = sum(sig0.05),
            sig0.01 = sum(sig0.01),
            total = n()) %>%
  mutate(per_sig0.05 = sig0.05 / total,
         per_sig0.01 = sig0.01 / total) %>%
  left_join(keys$continuous)

t_tests_b %>%
  separate(subzone, into = c("zone", "subzone")) %>% 
  mutate(sig0.05 = ifelse(p.value < 0.05, T, F),
         sig0.01 = ifelse(p.value < 0.01, T, F)) %>%
  group_by(variable) %>%
  summarize(sig0.05 = sum(sig0.05),
            sig0.01 = sum(sig0.01),
            total = n()) %>%
  mutate(per_sig0.05 = sig0.05 / total,
         per_sig0.01 = sig0.01 / total)

t_tests_scatter <- t_tests_zonal %>%
  filter(!variable %in% c("slope", "elev")) %>%
  mutate(axis_label = paste0(var_long, " (", unit, ")")) %>%
  ggplot(aes(x = per_sig0.01, y = axis_label)) +
  geom_boxplot(outlier.alpha = 0) +
  geom_jitter(aes(colour = zone), size = 2, width = 0) +
  #geom_text_repel(aes(label = zone), max.overlaps = 20) +
  labs(x = "% of Significant \n (P < 0.01) T-Tests",
       y = NULL,
       colour = NULL) +
  scale_colour_igv() +
  theme(legend.position = "bottom",
        legend.justification = c(0, 1),
        legend.margin = margin(t = 0, r = 0, b = 0, l = -69, unit = "pt")) +
  scale_x_continuous(labels = scales::percent, limits = c(0.5, 1)) +
  scale_y_discrete(labels = function(x) str_wrap(x, width = 15))

t_tests_scatter

ggsave(here::here("outputs", "t_tests_scatter.png"), plot = t_tests_scatter, device = "png",
       height = 5.5, width = 3)
  
```

Structure 3D Plot

```{r 3d-structure-subzone}
bec_cols <- pal_igv("default")(16)
structure <- read_csv(here::here("data", "structure_means.csv"))

for_arrows_subzone_full <- structure %>%
  dplyr::select(!ends_with("total_biomass")) %>%
  dplyr::select(protected, subzone, starts_with("mean")) %>%
  pivot_wider(names_from = protected, values_from = starts_with("mean")) %>%
  #%>%
  drop_na() #%>%
  #filter(zone != "SBPS") %>%

for_arrows_subzone <- structure %>% 
  dplyr::select(!ends_with("total_biomass")) %>%
  dplyr::select(!starts_with("sd") & !count) %>%
  mutate(across(starts_with("mean"), scale)) %>%
  pivot_wider(names_from = protected, values_from = starts_with("mean")) %>%
  mutate(elev_cv_dif = mean_elev_cv_protected - mean_elev_cv_unprotected,
         loreys_height_dif = mean_loreys_height_protected - mean_loreys_height_unprotected,
         percentage_first_returns_above_2m_dif = mean_percentage_first_returns_above_2m_protected - mean_percentage_first_returns_above_2m_unprotected) %>%
  dplyr::select(subzone, ends_with("dif")) %>%
  mutate(vector_length = sqrt((elev_cv_dif)^2 + (loreys_height_dif)^2 + (percentage_first_returns_above_2m_dif)^2)) %>%
  filter(vector_length > 1) %>% # this is the filter for which ones are selected
  left_join(for_arrows_subzone_full) %>%
  separate(subzone, into = c("zone", "subzone")) %>%
  mutate(id = row_number())

arrows_zones <- for_arrows_subzone %>%
  pull(zone) %>% 
  unique()
  
subzone_cols <- bec_cols[(keys$bec %>% pull(zone)) %in% arrows_zones]

for_arrows_subzone <- for_arrows_subzone %>% 
  mutate(label = paste(zone, subzone, sep = "-"))

# get the legend from this
for_arrows_subzone %>%
  ggplot(aes(x = id, y = vector_length, fill = label)) +
  geom_col() +
  labs(fill = "Zone-Subzone") +
  scale_fill_scico_d()

subzones_3d <- function() {
  #plotrgl()
  arrows3D(for_arrows_subzone$mean_elev_cv_protected, 
           for_arrows_subzone$mean_loreys_height_protected, 
           for_arrows_subzone$mean_percentage_first_returns_above_2m_protected,
           for_arrows_subzone$mean_elev_cv_unprotected, 
           for_arrows_subzone$mean_loreys_height_unprotected, 
           for_arrows_subzone$mean_percentage_first_returns_above_2m_unprotected,
           colvar = for_arrows_subzone$id,
           #col = subzone_cols,
           colkey = FALSE,
           lwd = 1, 
           d = 3, 
           #bty = "g",
           ticktype = "detailed",
           xlab = "\n Elevation Covariance",
           ylab = "\n Forest Height (m)",
           zlab = "\n Canopy Cover (%)",
           type = "cone"
           #xlim = c(-3, 3),
           #ylim = c(-3, 3),
           #zlim = c(-3, 3)
           )
  
  points3D(for_arrows_subzone$mean_elev_cv_protected, 
           for_arrows_subzone$mean_loreys_height_protected, 
           for_arrows_subzone$mean_percentage_first_returns_above_2m_protected,
           labels = for_arrows_subzone$label,
           colvar = for_arrows_subzone$id,
           #col = subzone_cols,
           add = T,
           colkey = F,
           type = "p",
           pch = 19,
           clab = "BEC Zone"#,
           #colkey = list(at = c(1:max(for_arrows_subzone$id)), side = 4,
           #             addlines = T, labels = unique(for_arrows_subzone$zone))
           )
}


subzones_3d_scatter <- plot_grid(subzones_3d) + bec_legend

subzones_3d_scatter

ggsave(here::here("outputs", "subzones_3d_scatter.png"),
       plot = subzones_3d_scatter, device = "png",
       height = 5.5,
       width = 10)

zones_img <- image_read(here::here("outputs", "subzones_3d_scatter.png")) 

zones_img %>% image_trim() %>% image_write(here::here("outputs", "subzones_3d_scatter.png"), format = "png")

```

```{r 3d-structure-zone}
# weighted means to account for varying subzone size

for_arrows_zone <- structure %>%
  separate(subzone, into = c("zone", "subzone")) %>%
  dplyr::select(!starts_with("sd")) %>%
  group_by(protected, zone) %>%
  summarize(mean_elev_cv = weighted.mean(mean_elev_cv, count),
            mean_percentage_first_returns_above_2m = weighted.mean(mean_percentage_first_returns_above_2m, count),
            mean_loreys_height = weighted.mean(mean_loreys_height, count),
            mean_total_biomass = weighted.mean(mean_total_biomass, count)) %>%
  pivot_wider(names_from = protected, values_from = starts_with("mean")) %>%
  #filter(zone != "SBPS") %>%
  mutate(id = row_number())

structure %>%
  separate(subzone, into = c("zone", "subzone")) %>%
  dplyr::select(!starts_with("sd")) %>%
  group_by(protected, zone) %>%
  summarize(mean_elev_cv = weighted.mean(mean_elev_cv, count),
            mean_percentage_first_returns_above_2m = weighted.mean(mean_percentage_first_returns_above_2m, count),
            mean_loreys_height = weighted.mean(mean_loreys_height, count),
            mean_total_biomass = weighted.mean(mean_total_biomass, count)) %>%
  mutate(across(starts_with("mean"), scale)) %>%
  pivot_wider(names_from = protected, values_from = starts_with("mean")) %>%
  mutate(elev_cv_dif = mean_elev_cv_protected - mean_elev_cv_unprotected,
         loreys_height_dif = mean_loreys_height_protected - mean_loreys_height_unprotected,
         percentage_first_returns_above_2m_dif = mean_percentage_first_returns_above_2m_protected - mean_percentage_first_returns_above_2m_unprotected) %>%
  dplyr::select(zone, ends_with("dif")) %>%
  mutate(vector_length = sqrt((elev_cv_dif)^2 + (loreys_height_dif)^2 + (percentage_first_returns_above_2m_dif)^2)) %>%
  arrange(desc(vector_length))
  

zones_3d <- function() {

  arrows3D(for_arrows_zone$mean_elev_cv_protected, 
           for_arrows_zone$mean_loreys_height_protected, 
           for_arrows_zone$mean_percentage_first_returns_above_2m_protected,
           for_arrows_zone$mean_elev_cv_unprotected, 
           for_arrows_zone$mean_loreys_height_unprotected, 
           for_arrows_zone$mean_percentage_first_returns_above_2m_unprotected,
           col = bec_cols,
           lwd = 1, 
           d = 3, 
           #bty = "g",
           ticktype = "detailed",
           xlab = "\n Elevation Covariance",
           ylab = "\n Forest Height (m)",
           zlab = "\n Canopy Cover (%)",
           type = "cone")
  
  points3D(for_arrows_zone$mean_elev_cv_protected, 
           for_arrows_zone$mean_loreys_height_protected, 
           for_arrows_zone$mean_percentage_first_returns_above_2m_protected,
           labels = for_arrows_zone$zone,
           colvar = for_arrows_zone$id,
           col = bec_cols, 
           add = T,
           colkey = F,
           type = "p",
           pch = 19,
           clab = "BEC Zone"
           #colkey = list(at = c(1:max(for_arrows_zone$id)), side = 4,
          #               addlines = T, labels = for_arrows_zone$zone)
           )
}

zones_3d_scatter <- plot_grid(zones_3d) + bec_legend #& plot_annotation(theme = theme(plot.margin = margin(0, 5, 0, 5, "cm")))

ggsave(here::here("outputs", "zones_3d_scatter.png"),
       plot = zones_3d_scatter, device = "png")

# cropping because i couldn't figure it out in GGPLOT lol

zones_img <- image_read(here::here("outputs", "zones_3d_scatter.png")) 

zones_img %>% image_trim() %>% image_write(here::here("outputs", "zones_3d_scatter.png"), format = "png")

```

Latitude Disturbance

```{r lat-dist}
dist_colours <- c("#FF0000", "#38a800", "#F0F0F0")

bc_bounds <- read_sf(here::here("data", "shapefiles", "bc_boundary.shp"))

lat_bbox <- bc_bounds %>% st_transform(4326) %>% st_bbox()

lat_range <- lat_bbox$ymax - lat_bbox$ymin

lat_agg <- 0.5
  
min_lat <- lat_dist %>%
  pull(latitude) %>%
  min()

max_lat <- lat_dist %>%
  pull(latitude) %>%
  max()
    
lat_df <- lat_dist %>% 
  group_by(protected, latitude, Change_Attribution) %>%
  summarize(count = sum(count)) %>%
  mutate(latitude = latitude - min_lat,
         per_lat = latitude / max_lat,
         lat_real = (per_lat * lat_range) + lat_bbox$ymin,
         lat_real = lat_real %/% lat_agg * lat_agg) %>%
  mutate(Change_Attribution = case_when(Change_Attribution == 1 ~ 1,
                                        Change_Attribution == 2 ~ 2,
                                        TRUE ~ 0)) %>%
  group_by(protected, lat_real, Change_Attribution) %>%
  summarize(count = sum(count)) %>%
  mutate(perc = count / sum(count)) %>%
  left_join(keys$disturbance, by = c("Change_Attribution" = "class_val"))

lat_plot <- lat_df %>%
  ggplot(aes(x = lat_real, y = perc, fill = class)) +
  geom_area() +
  facet_wrap(~str_to_title(protected)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 5L)) +
  coord_flip(ylim = c(.75, 1)) +
  labs(x = expression("Latitude " ( degree~N)),
       y = "% Disturbed",
       fill = NULL) +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = dist_colours) +
  scale_x_continuous(breaks = seq(48, 60, 3))

lat_plot_prot <- lat_df %>%
  filter(protected == "protected") %>%
  ggplot(aes(x = lat_real, y = perc, fill = class)) +
  geom_area() +
  facet_wrap(~str_to_title(protected)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 5L)) +
  coord_flip(ylim = c(.75, 1)) +
  labs(x = expression("Latitude " ( degree~N)),
       y = "% Disturbed",
       fill = NULL) +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = dist_colours) +
  scale_x_continuous(breaks = seq(48, 60, 3))

lat_plot_unprot <- lat_df %>%
  filter(protected == "unprotected") %>%
  ggplot(aes(x = lat_real, y = perc, fill = class)) +
  geom_area() +
  facet_wrap(~str_to_title(protected)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 5L)) +
  coord_flip(ylim = c(.75, 1)) +
  labs(x = NULL,
       y = "% Disturbed",
       fill = NULL) +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = dist_colours) +
  scale_x_continuous(breaks = seq(48, 60, 3)) +
  theme(axis.text.y = element_blank())

lat_prot <- lat_df %>%
  group_by(lat_real, protected) %>%
  summarize(count = sum(count)) %>%
  mutate(perc = count / sum(count)) %>%
  filter(protected == "protected") %>%
  mutate(label = "% Protected") %>%
  ggplot(aes(x = lat_real, y = perc)) +
  #geom_smooth() +
  #geom_point() +
  geom_area() +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
  labs(x = NULL,
       y = NULL,
       fill = NULL) +
  theme(axis.text.y = element_blank(),
        legend.position = c(0.7, 0.85)) +
  coord_flip() +
  scale_x_continuous(breaks = seq(48, 60, 3)) +
  facet_wrap(~label)
  

dist_lat_hist <- (lat_plot_prot + lat_plot_unprot + lat_prot) +
  plot_annotation(tag_levels = "a") +
  plot_layout(nrow = 1, 
              guides = "collect",
              widths = c(1, 1, 1)) &
  theme(legend.position = "bottom")

ggsave(here::here("outputs", "latitude_disturbance_plot.png"),
       plot = dist_lat_hist, device = "png",
       height = 5.5,
       width = 8.5)
```

```{r}
lat_df <- lat_dist %>% 
  separate(subzone, into = c("zone", "subzone")) %>%
  group_by(zone, protected, latitude, Change_Attribution) %>%
  summarize(count = sum(count)) %>%
  mutate(latitude = latitude - min_lat,
         per_lat = latitude / max_lat,
         lat_real = (per_lat * lat_range) + lat_bbox$ymin,
         lat_real = lat_real %/% lat_agg * lat_agg) %>%
  mutate(Change_Attribution = case_when(Change_Attribution == 1 ~ 1,
                                        Change_Attribution == 2 ~ 2,
                                        TRUE ~ 0)) %>%
  group_by(zone, protected, lat_real, Change_Attribution) %>%
  summarize(count = sum(count)) %>%
  mutate(perc = count / sum(count)) %>%
  left_join(keys$disturbance, by = c("Change_Attribution" = "class_val"))

lat_output <- lat_dist %>%
  ungroup() %>%
  separate(subzone, into = c("zone", "subzone")) %>%
  group_by(zone, protected, latitude, Change_Attribution) %>%
  summarize(count = sum(count)) %>%
  mutate(latitude = latitude - min_lat,
         per_lat = latitude / max_lat,
         lat_real = (per_lat * lat_range) + lat_bbox$ymin,
         lat_real = lat_real %/% lat_agg * lat_agg) %>%
  mutate(Change_Attribution = case_when(Change_Attribution == 1 ~ 1,
                                        Change_Attribution == 2 ~ 2,
                                        TRUE ~ 0)) %>%
  group_by(zone, protected, lat_real, Change_Attribution) %>%
  summarize(count = sum(count)) %>%
  mutate(perc = count / sum(count)) %>%
  left_join(keys$disturbance, by = c("Change_Attribution" = "class_val"))

lat_output %>%
  ggplot(aes(x = lat_real, y = perc, fill = class)) +
  geom_area() +
  facet_grid(cols ~ vars(protected),
             rows ~ vars(zone)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 5L)) +
  coord_flip(ylim = c(.75, 1)) +
  labs(x = expression("Latitude " ( degree~N)),
       y = "% Disturbed",
       fill = NULL) +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = dist_colours) +
  scale_x_continuous(breaks = seq(48, 60, 3))

lat_output %>%
  ggplot(aes(x = lat_real, y = perc, fill = class)) +
  geom_area() +
  facet_grid(rows = vars(zone),
             cols = vars(protected),
             scales = "free_x") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 5L)) +
  coord_flip(ylim = c(.75, 1)) +
  labs(x = expression("Latitude " ( degree~N)),
       y = "% Disturbed",
       fill = NULL) +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = dist_colours)

ggsave(here::here("outputs", "lat_dist_zone.png"), device = "png",
       width = 10, height = 45)
```



Table 1: BEC info

```{r bec-table}
bec <- read_sf(here::here("data", "shapefiles", "bec_subzones.shp"))

bec_df <- bec %>% 
  mutate(area = st_area(.)) %>%
  st_drop_geometry()

num_subzones <- bec_df %>%
  group_by(zone, subzone) %>%
  summarize(area = sum(area)) %>%
  group_by(zone) %>%
  summarize(n_subzones = n())

bec_area <- bec_df %>%
  group_by(zone) %>%
  summarize(area = sum(area))

bec_ppa <- read_sf(here::here("data", "shapefiles", "bc_ppa_bec_hres.shp"))

bec_ppa_df <- bec_ppa %>%
  mutate(area = st_area(.),
         protected = ifelse(!is.na(IUCN_CA), T, F)) %>%
  st_drop_geometry()

bec_perc <- bec_ppa_df %>% 
  group_by(zone, protected) %>% 
  summarize(area = sum(area)) %>%
  mutate(perc = area / sum(area)) %>%
  filter(protected == T) %>%
  drop_na() %>%
  dplyr::select(zone, perc_protected = perc)

bec_table <- keys$bec %>% 
  left_join(num_subzones) %>%
  left_join(bec_area) %>%
  left_join(bec_perc) %>%
  mutate(area = as.integer(as.numeric(area / 10000)), # converts to ha?
         perc_protected = paste0(round(as.numeric(perc_protected * 100), 1), "%"))

write_csv(bec_table, here::here("outputs", "bec_table.csv"))
```

Copy of the cell that generates the table in the paper

```{r}
bec_table <- read_csv(here::here("outputs", "bec_table.csv"))

bec_table

bec_ft <- flextable(bec_table) %>%
  set_header_labels(zone = "Zone",
                    zone_nm = "Zone Name",
                    n_subzones = "# of Subzones",
                    area = "Area (ha)",
                    perc_protected = "# Protected") %>%
  set_caption("Number of subzones, total area, and percent protected by BEC Zone",
              autonum = run_autonum(seq_id = "tab", bkm = "bec-table"),
              style = "Table Caption") %>%
  set_table_properties(layout = "autofit")

bec_ft
```

```{r subzones-missing-counterparts}
vlce %>% 
  left_join(keys$vlce, by = c("vlce" = "class_val")) %>%
  mutate(other = ifelse(forest == "Other", "Other", "Not-other")) %>%
  group_by(subzone, protected, other) %>%
  summarize(count = sum(n)) %>%
  mutate(perc = count / sum(count)) %>%
  dplyr::select(!count) %>%
  pivot_wider(names_from = protected, values_from = perc) %>%
  filter(other == "Other") %>%
  filter(is.nan(protected) | is.nan(unprotected))
```

```{r}
for_flex <- structure %>%
  separate(subzone, into = c("zone", "subzone")) %>%
  dplyr::select(!starts_with("sd")) %>%
  group_by(protected, zone) %>%
  summarize(mean_elev_cv = weighted.mean(mean_elev_cv, count),
            mean_percentage_first_returns_above_2m = weighted.mean(mean_percentage_first_returns_above_2m, count),
            mean_loreys_height = weighted.mean(mean_loreys_height, count),
            mean_total_biomass = weighted.mean(mean_total_biomass, count)) %>%
  dplyr::select(!mean_total_biomass) %>%
  pivot_wider(names_from = protected, values_from = starts_with("mean")) %>%
  mutate(mean_elev_change = (1 - (mean_elev_cv_protected / mean_elev_cv_unprotected)) * 100,
         mean_per_change = (1 - (mean_percentage_first_returns_above_2m_protected / mean_percentage_first_returns_above_2m_unprotected)) * 100,
         mean_lor_change = (1 - (mean_loreys_height_protected / mean_loreys_height_unprotected)) * 100) %>%
  dplyr::select(zone,
         mean_elev_cv_protected,
         mean_elev_cv_unprotected,
         mean_elev_change,
         mean_percentage_first_returns_above_2m_protected,
         mean_percentage_first_returns_above_2m_unprotected,
         mean_per_change,
         mean_loreys_height_protected,
         mean_loreys_height_unprotected,
         mean_lor_change)

zones <- for_flex %>%
  pull(zone)

highlighter <- for_flex %>%
  dplyr::select(zone, ends_with("change")) %>%
  pivot_longer(cols = ends_with("change")) %>%
  separate(name, into = c("mean", "var", "change")) %>%
  filter(abs(value) >= 5) %>%
  group_split(var) %>%
  set_names("elev", "lor", "per") %>%
  map(pull, zone)
  

index_highlight <- function(valid_zone) {
  index_out <- zones %in% valid_zone
  
  return(which(index_out))
}

highlight_indexes <- map(highlighter, index_highlight)

for_flex %>%
  mutate(across(ends_with("protected"), round, digits = 2),
         across(ends_with("change"), round, digits = 2),
         across(ends_with("change"), paste0, "%"),
         across(starts_with("mean_perc"), paste0, "%")) %>%
  flextable() %>%
  add_header_row(values = c("Zone", "Elevation Covariance", "Canopy Cover (%)", "Canopy Height (m)"),
                 colwidths = c(1, 3, 3, 3)) %>%
  set_header_labels(
    mean_elev_cv_protected = "Protected",
    mean_elev_cv_unprotected = "Unprotected",
    mean_elev_change = "% Change",
    mean_percentage_first_returns_above_2m_protected = "Protected",
    mean_percentage_first_returns_above_2m_unprotected = "Unprotected",
    mean_per_change = "% Change",
    mean_loreys_height_protected = "Protected",
    mean_loreys_height_unprotected = "Unprotected",
    mean_lor_change = "% Change",
    zone = "Zone") %>%
  theme_booktabs(bold_header = T) %>%
  align(part = "all", align = "center") %>%
  border_inner_v(border = fp_border(color = "grey"), part = "all") %>%
  merge_v(part = "header") %>%
  colformat_num() %>%
  autofit() %>%
  bg(i = highlight_indexes$elev,
     j = 2:4,
     bg = "grey") %>%
  bg(i = highlight_indexes$per,
     j = 5:7,
     bg = "grey") %>%
  bg(i = highlight_indexes$lor,
     j = 8:10,
     bg = "grey")
```

Some plots Txomin suggested. May be used, leaving in 05_vis

```{r matrix}
structure %>%
  separate(subzone, into = c("zone", "subzone")) %>%
  dplyr::select(!starts_with("sd")) %>%
  group_by(protected, zone) %>%
  summarize(mean_elev_cv = weighted.mean(mean_elev_cv, count),
            mean_percentage_first_returns_above_2m = weighted.mean(mean_percentage_first_returns_above_2m, count),
            mean_loreys_height = weighted.mean(mean_loreys_height, count),
            mean_total_biomass = weighted.mean(mean_total_biomass, count)) %>%
  mutate(across(starts_with("mean"), scale)) %>%
  pivot_longer(starts_with("mean")) %>%
  ggplot(aes(x = zone, y = str_to_title(protected), fill = value)) +
  geom_tile() +
  facet_wrap(~name) +
  labs(x = NULL,
       y = NULL,
       fill = "Z-Score") +
  scale_fill_scico(palette = "vik") +
  coord_flip()

z_score_plot <- structure %>%
  separate(subzone, into = c("zone", "subzone")) %>%
  dplyr::select(!starts_with("sd")) %>%
  group_by(protected, zone) %>%
  summarize(mean_elev_cv = weighted.mean(mean_elev_cv, count),
            mean_percentage_first_returns_above_2m = weighted.mean(mean_percentage_first_returns_above_2m, count),
            mean_loreys_height = weighted.mean(mean_loreys_height, count),
            mean_total_biomass = weighted.mean(mean_total_biomass, count)) %>%
  mutate(across(starts_with("mean"), scale)) %>%
  pivot_longer(starts_with("mean"), names_to = "var") %>%
  pivot_wider(names_from = protected, values_from = value) %>%
  mutate(difference = protected - unprotected) %>%
  pivot_longer(protected:difference, names_to = "protected") %>%
  mutate(var = str_remove(var, "mean_")) %>%
  left_join(keys$continuous, by = c("var" = "variable")) %>%
  ggplot(aes(x = value, y = fct_rev(zone), fill = str_to_title(protected))) +
  geom_col(position = "dodge") +
  facet_wrap(~str_replace(var_long, " ", "\n")) + #replaces  with newline so it facets and doesn't cut off due to the small x width (3 inch)
  labs(x = "Z-Score",
       y = NULL,
       fill = NULL) +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(ncol = 1)) +
  scale_fill_manual(values = c("#18db02" , "grey50", "grey70"))

z_score_plot

ggsave(here::here("outputs", "fstruct_zscores.png"), plot = z_score_plot, device = "png",
       height = 8, width = 3)

structure %>%
  separate(subzone, into = c("zone", "subzone")) %>%
  dplyr::select(!starts_with("sd")) %>%
  group_by(protected, zone) %>%
  summarize(mean_elev_cv = weighted.mean(mean_elev_cv, count),
            mean_percentage_first_returns_above_2m = weighted.mean(mean_percentage_first_returns_above_2m, count),
            mean_loreys_height = weighted.mean(mean_loreys_height, count),
            mean_total_biomass = weighted.mean(mean_total_biomass, count)) %>%
  #mutate(across(starts_with("mean"), scale)) %>%
  pivot_longer(starts_with("mean"), names_to = "var") %>%
  pivot_wider(names_from = protected, values_from = value) %>%
  mutate(difference = protected - unprotected) %>%
  pivot_longer(protected:difference, names_to = "protected") %>%
  mutate(var = str_remove(var, "mean_")) %>%
  left_join(keys$continuous, by = c("var" = "variable")) %>%
  ggplot(aes(x = value, y = fct_rev(zone), fill = str_to_title(protected))) +
  geom_col(position = "dodge") +
  facet_wrap(~var_long, scales = "free_x") +
  labs(x = "Z-Score",
       y = "Zone",
       fill = NULL) +
  theme(legend.position = "bottom") +
  scale_fill_brewer(palette = "Dark2")

```


map making (why did i do this in R)

```{r map}
library(rworldxtra)

bec_sf <- read_sf(here::here("data", "shapefiles", "bec_zones.shp"))

bec_zones <- aggregate(x = bec_sf$are_sqm, by = list(bec_sf$zone),
                       FUN = sum, na.rm = TRUE,
                       do.union = T)

bec_zones = bec_sf %>% group_by(zone) %>%
  summarize(area = sum(are_sqm, na.rm = TRUE))

bc_bounds <- read_sf(here::here("data", "shapefiles", "bc_boundary.shp"))

bec_zones_terr <- bec_zones %>%
  st_intersection(bc_bounds)

parks <- read_sf(here::here("data", "shapefiles", "bc_ppa.shp"))

provinces <- getData(country = "Canada", level = 1)

prov_sf <- provinces %>%
  st_as_sf() %>%
  st_transform(3005)

bbox <- bc_bounds %>%
  st_bbox()

data("countriesHigh")

#rworldxtra

usa <- countriesHigh %>%
  st_as_sf() %>%
  st_transform(3005) %>%
  filter(ADMIN == "United States of America")

countries <- countriesHigh %>%
  st_as_sf() %>%
  st_transform(3005)

na_bbox <- countries %>%
  filter(ADMIN == "United States of America" |
           ADMIN == "Canada") %>%
  st_bbox()

cad_bbox <- prov_sf %>%
  st_transform(3348) %>%
  st_bbox()

bec_zones_terr <- bec_zones_terr %>%
  left_join(keys$bec)

bec_map <- ggplot() +
  geom_sf(data = bec_zones_terr, aes(fill = fct_reorder(zone_nm, zone)), colour = NA) +
  geom_sf(data = parks, aes(linetype = "Protected Areas"), fill = "Green") +
  #geom_sf(data = bc_bounds, fill = NA, size = 0.5) +
  geom_sf(data = usa) +
  geom_sf(data = prov_sf %>% filter(NAME_1 != "British Columbia"), fill = "grey70") +
  scale_linetype_manual(labels = "Protected Areas", values = "solid") +
  scale_fill_igv() +
  labs(fill = NULL,
       linetype = NULL) +
  coord_sf(xlim = c(bbox$xmin - 300000, bbox$xmax), 
           ylim = c(bbox$ymin, bbox$ymax),
           crs = 3005,
           datum = 4326) +
  theme_grey() +
  theme(panel.background = element_rect(fill = "lightblue"),
        panel.border = element_rect(colour = "black", fill = NA),
        legend.position = c(.15, .4),
        legend.background = element_rect(fill = NA),
        legend.box.background = element_rect(fill = "white"))

#ggsave(here::here("outputs", "bec_map_no_inset.png"), bec_map, device = "png")

countries_map <- countriesHigh %>%
  st_as_sf() %>%
  filter(ADMIN == "United States of America" |
         ADMIN == "Greenland") %>%
  st_transform(3348)
  

inset = ggplot() +
  geom_sf(data = countries_map) +
  geom_sf(data = prov_sf %>% st_transform(3348) %>% filter(NAME_1 != "British Columbia"), fill = "grey70") +
  geom_sf(data = prov_sf %>% st_transform(3348) %>% filter(NAME_1 == "British Columbia"), fill = "grey50") +
  geom_sf(data = bbox %>% st_as_sfc %>% st_transform(3348), fill = NA, colour = "red", size = 1.2) +
  coord_sf(xlim = c(cad_bbox$xmin - 300000, cad_bbox$xmax),
           ylim = c(cad_bbox$ymin, cad_bbox$ymax)) +
  theme_void() +
  theme(panel.border = element_rect(colour = "black"),
        panel.background = element_rect(fill = "lightblue"))

map_inset <- ggdraw(bec_map) +
  draw_plot(inset, x = 0.7, y = 0.7, width = .28, height = .28)

  
ggsave(here::here("outputs", "bec_map.png"), plot = map_inset, device = "png", height = 7, width = 10)
```

Garabaldi Data Map

```{r gar-map}
bec <- read_sf(here::here("data", "shapefiles", "bec_zones.shp"))
bc_ppa <- read_sf(here::here("data", "shapefiles", "bc_ppa.shp"))

gar_sf <- bc_ppa %>% 
  filter(NAME_E == "Garibaldi Park") %>%
  st_buffer(10000)

plot(gar_sf %>% st_geometry())

gar_vect <- gar_sf %>%
  vect()

gar_bec <- st_intersection(gar_sf, bec)

in_rasters <- c(here::here("G:/", "Merged", "vlce", "BC-vlce-2015.tif"),
                here::here("G:/", "Merged", "Change_Attribution", "BC-Change_Attribution.tif"),
                here::here("G:/", "Merged", "elevation", "BC-elev.tif"),
                here::here("G:/", "Merged", "masked_structure", "elev_cv", "masked-elev_cv-2015.tif"),
                here::here("G:/", "Merged", "masked_structure", "loreys_height", "masked-loreys_height-2015.tif"),
                here::here("G:/", "Merged", "masked_structure", "total_biomass", "masked-total_biomass-2015.tif"),
                here::here("G:/", "Merged", "masked_structure", "percentage_first_returns_above_2m", "masked-percentage_first_returns_above_2m-2015.tif"))

struct_vars <- c("elev_cv",
                 "loreys_height",
                 "total_biomass",
                 "percentage_first_returns_above_2m")

clip_gar <- function(filepath) {
  
  var <- str_split(filepath, "/")[[1]] %>%
    tail(2) %>%
    head(1)
  print(var)

  my_rast <- rast(filepath)
  
  my_rast <- my_rast %>%
    crop(gar_vect) %>%
    terra::mask(gar_vect)
  
  if (var %in% struct_vars) {
    divider <- keys$continuous %>%
      filter(variable == var) %>%
      pull(divide_by)
    print(divider)
    
    my_rast = my_rast / divider
  }
  
  if (var == "elev_cv") {
    # removes 5.912 outlier in the elev_cv dataset
    
    my_rast <- subst(my_rast, 5.912, NA)
  }
  
  writeRaster(my_rast,
              filename = here::here("data", "gar_maps", paste0("gar-", var, ".tif")),
              overwrite = T)
  
  return(my_rast)
}

gar_rasts <- map(in_rasters, clip_gar)

gar_slope <- terrain(gar_rasts[[3]], v = "slope", unit = "radians")
gar_asp <- terrain(gar_rasts[[3]], v = "aspect", unit = "radians")

gar_shade <- shade(gar_slope, gar_asp)

terra::writeRaster(gar_shade, 
                   here::here("data", "gar_maps", "hillshade.tif"),
                   overwrite = T)

terra::writeVector(gar_bec %>% vect(), 
                   here::here("data", "gar_maps", "gar_bec.shp"),
                   overwrite = T)
```

```{r quantiles}
data <- readr::read_csv(here::here("data", "all_vars", "sampled", "BAFA_un-2015.csv"))

p = c(0.025, 0.975)

data %>%
  group_by(protected) %>%
  summarize(across(elev_cv:total_biomass, quantile, probs = p)
```

